# Fractions

```{r setup, include = FALSE}
source("R/setup.R")
```

```{r}
#| code-summary: Setup

library(tidyverse)
theme_set(theme_minimal(base_size = 15))

# devtools::install_github("scienceverse/papercheck")
library(papercheck)
```

Papercheck is an R package I'm working on with Daniel Lakens and colleagues. The purpose is to help researchers assess best practices in methodology and reporting. It can also be used to make meta-scientific work scriptable and easier.

The package comes with a built-in dataset of 250 open-access papers from Psychological Science.

You can use the `search_text()` function to search any sentences that contain the word "badge".

```{r}
badges <- psychsci |>
  search_text("badge")
```

Check all the permutations of the badge statements. 

```{r}
count(badges, text, sort = TRUE)
```

It looks like the relevant statements all contain the word "received".

```{r}
badges <- psychsci |>
  search_text("badge") |>
  search_text("received")
```

```{r}
count(badges, text, sort = TRUE)
```


Use `grepl()` to determine if these sentences contain the text "open data", "open materials" and/or "preregistration".

```{r}
badges <- badges |>
  mutate(data = grepl("open data", text, ignore.case = TRUE),
         materials = grepl("open materials", text, ignore.case = TRUE),
         prereg = grepl("preregistration", text, ignore.case = TRUE),
         ) |>
  select(id, data, materials, prereg)
```

Now you can use the `info_table()` function to get information about each paper's submission and acceptance dates. Then join this to the badge data and set NA values to FALSE. 

```{r}
all_papers <- psychsci |>
  info_table(c("submission")) |>
  separate(submission, c("received", "accepted"), sep = "; ") |>
  left_join(badges, by = "id") |>
  replace_na(replace = list(data = FALSE, 
                            materials = FALSE, 
                            prereg = FALSE))
```

Sort out the dates...

```{r}
m <- gregexpr("\\d{1,2}/\\d{1,2}/\\d{2}", all_papers$received)
all_papers$year_received <- regmatches(all_papers$received, m) |> 
  lapply(mdy) |>
  sapply(\(x) ifelse(length(x), year(x), NA))

m <- gregexpr("\\d{1,2}/\\d{1,2}/\\d{2}", all_papers$accepted)
all_papers$year_accepted<- regmatches(all_papers$accepted, m) |> 
  lapply(mdy) |>
  sapply(\(x) ifelse(length(x), year(x), NA))
```

Calculate the fraction of papers each year with open data, materials and preregistration.


```{r}
by_year <- all_papers |>
  fill(year_received, year_accepted) |>
  pivot_longer(data:prereg) |>
  summarise(frac = mean(value), .by = c(year_received, name))
```



```{r}
frac <- c(0, 1/10, 1/5, 1/4, 1/3, 2/5, 1/2, 3/5,  2/3, 3/4, 4/5, 9/10, 1)
label <- c("0", "1/10", "1/5", "1/4", "1/3", "2/5", "1/2", "3/5", "2/3", "3/4", "4/5", "9/10", "1")

ggplot(by_year, aes(x = year_received, y = frac, colour = name)) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_continuous("Year Received", breaks = 2012:2024) +
  scale_y_continuous("Fraction of Papers", limits = c(0, 1),
                     breaks = frac, labels = label) +
  scale_colour_manual("Badge:", values = c("firebrick", "darkgreen", "dodgerblue3")) +
  labs(title = "Open Practice Badges",
       subtitle = "250 Open Access Psychologial Science Papers") +
  theme(legend.position = c(.2, .8))
```

Something weird was going on in 2022. Let's figure it out tomorrow.

```{r}
#| echo: false
ggsave("images/day1.png", width = 8, height = 8)
```

